---
title: "formative.rmd"
author: "kimp"
date: "10/5/2017"
output: html_document
---

# Load Dyplyr and Tidyr

```{r}
library(dplyr)
library(tidyr)
```


# Upload the MITx/Harvardx dataset to R as a dataframe named D1.

```{r}
D1 <- read.csv("HMXPC13_DI_v2_5-14-14.csv")
head(D1)
```

# What does the DI suffix that is appeneded to two of the variable names mean?

There is a note that details the csv file: 
"MITx and HarvardX, 2014, "HarvardX-MITx Person-Course Academic Year 2013 De-Identified dataset, version 2.0", doi:10.7910/DVN/26147, Harvard Dataverse, V10; HMXPC13_DI_v2_5-14-14.csv [fileName]"
(https://dataverse.harvard.edu/file.xhtml?fileId=2468954&version=RELEASED&version=.0)

The DI suffix appends 4 varaible names (userid, final_cc_cname, start_time, and last_event). The DI should stand for de-identified data.

# How many people enrolled in HarvardX/CS50x recieved certification in 2012? (Please inlcude your code in your markdown document)

```{r}
x <- filter(D1, course_id == 'HarvardX/CS50x/2012')
head(x)
count(x, certified)
```

x <- filter(D1, course_id == 'HarvardX/CS50x/2012')
head(x)
count(x, certified)

1287 people enrolled in HarvardxCS50x and received certification in 2012.

# How many people registered for HarvardX/CS50x from Columbia on October 15, 2012? (Please inlcude your code in your markdown document)

```{r}
y <- filter(D1, final_cc_cname_DI == 'Colombia', last_event_DI == '2012-10-15')
print(y)
count(y, certified)
```

y <- filter(D1, final_cc_cname_DI == 'Colombia', last_event_DI == '2012-10-15')
print(y)
count(y, certified)

In the printed data frame, there are 0 students who were registered for HarvardX/CS50x from Colombia on October 15, 2012.

# Create a new data frame (D2) that includes student ID and registration date for all students from a country of your choice (Please inlcude your code in your markdown document)

```{r}
z <- filter(D1, final_cc_cname_DI == 'United States')
head(z)
D2 <- select(z, userid_DI, start_time_DI)
head(D2)
```

z <- filter(D1, final_cc_cname_DI == 'United States')
head(z)
D2 <- select(z, userid_DI, start_time_DI)
head(D2)

# Create a new data frame (D3) in which the column names are the registration dates, with one column containing the student ids and all other cells indicating a 1 if the student registered on that date or an NA if they did not (hint: you will need to create a new variable before you change the structure of your data frame)


```{r}
dup <- select(D1, userid_DI, start_time_DI, registered)
dup <- mutate(dup, row = row_number())
D3 <- spread(dup, start_time_DI, registered)
head(D3)
```

# Now make the row names of D3 the student ids on your new data frame and remove the user id column from your data frame

```{r}
names(D3)[2]<-"studentid"
D3 <- select(D3, -userid_DI)
head(D3)
```

# Convert your data frame to a matrix (M1)

```{r}
M1 <- data.matrix(D3)
type(M1)
M1
```

# Create a new matrix which is the transposed version of your matrix (M2)

```{r}
M2 <- t(M1)
M2
```

# Create a third matrix (M3) that is a student by student matrix where the cells in the matrix represent if two studentes registered on the same day (This will take a lot of processing power, you may not be able to complete it if you have a large number of students)

```{r}
M3 <- intersect(M1[1,], M1[2,])

```

# Convert the diagonal of your matrix into NAs

```{r}
diag(M3) <- NA
```

